Local Search Approximation Algorithms for Partitional and Hierarchical Clustering

### 2015-05-08 tex/project.tex
Some implemenation and results work

We will compare the above algorithms against $k$-means, $k$-medoids, PAM, affinity propagation, and Shi and Meila's spectral algorithm. All these methods except $k$-medoids and PAM are implemented in python's scikit-learn package. For comparisons against affinity propagation, we will first observe how many clusters the algorithm outputs and use that as input for our proposed algorithms. We implement the above local search algorithm for $p=2, 3,4$ so the computation time is reasonable.

Given a true clustering, we will use the adjusted mutual information (AMI) measure to score the accuracy of each method. AMI measures the agreement of two assignments and is normalized to reflect chance so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation \cite{Vinh}. For each data set, we will also report the $k$-median objective function value for that clustering, the guaranteed approximation ratio, and the computation time.

The data sets we chose to use for comparison were

NEED TO DECIDE:
OR library $p$-median.

\section{Hierarchical Clustering Approximation Algorithms}
Again, suppose we have a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$. In the hierarchical setting, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$. Therefore, for any $i \in S$, $a^{-1}(i)$ will be all points in the cluster centered at $i$. We define
\[ \mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i)) . \]

We can test this hierarchical framework for both the multiswap and supermodular local search algorithms above.

@@

@@

\section{Data Sets}
In order to empirically test these clustering methods, we used four different data sets.  
\begin{description}
\item[Generated Gaussian Data] We wanted to test our clustering algorithms on data that had data nicely clustered around $k$ centers. So, we generated synthetic $k$-center Gaussian data algorithmically and used this for testing.
\item[$P$-median]
\item[UCI Iris Dataset] This is one of the most used data sets in pattern recognition (\cite{Iris}).  It contains data on 150 instances of flowers, each of with belongs to one of three species of iris.  One of the species is linearly separable from the others, but the other two are not.  In total, there are 150 instances and 4 attributes.
\item[UCI Soybean Dataset] This is another small, real data set from UCI.  This, too, is widely used in machine learning.  It has 47 instances, with 35 attributes.
\end{description}

\
\label{submission}

>>>>

\subsection{Implementation and Results}

We implement the above local search algorithm for $p=1, 2, 3$ so the computation time is reasonable, and compare the above algorithms against $k$-medoids and Shi and Meila's spectral algorithm in time, AMI, and approximation ratio.  Note that for $p=1$, the local search algorithm is equivalent to PAM.  In order to calculate approximation ratios, Gurobi was used to compute the global optimum for each data set.

To achieve reasonable computation times for the local search and supermodular function minimization algorithms, at each iteration we first consider nearby swaps, as well as placing

We also considered a hybrid heuristic algorithm where we ran $k$-medoids on the result of the local search algorithm.  This is motivated by the fact that, similar to Lloyd's algorithm, $k$-medoids will quickly converge to local optima.  So

When information on true labels is available, we use adjusted mutual information (AMI) measure to score accuracy of predicted clusters. AMI measures the agreement of two assignments and is normalized to reflect chance so that a perfect labeling receives a score of 1 and an independent random labeling receives a score of 0 in expectation.  See \cite{Vinh} for more information. For each data set and each clustering method, we also report objective function values, the approximation ratio, and the computation time.

The data sets tested against were:
\begin{description}
\item[Gaussian] Synthetic $k$-center Gaussian data, generated in Python.  Add info on std deviation, size, etc.
\item[P-median] ADD CITATION OR Library p-median data sets
\item[UCI Iris] \cite{Iris} Contains 4 attributes on 150 instances of flowers, each of with belongs to one of three species of iris.  One of the species is linearly separable from the others, but the other two are not.  One of the most used data sets in pattern recognition.
\item[UCI Soybean (Small)] \cite{Soybean} 47 instances, with 35 attributes data set of soybean disease from UCI.
\end{description}

@@

Tests were conducted using Python 2.7.9 on a 64-bit Windows platform, on an 8 core ....

@@

Comment on results here.

\section{Hierarchical Clustering Approximation Algorithms}
Suppose we have a set $X$ and a distance metric $d(i,j)$ for all $i,j \in X$. In the hierarchical setting, we will represent a solution to the $k$-median problem as a pair $(S, a)$ where $S \subseteq X$ such that $|S| \leq k$ and $a$ is an assignment function that maps each $i \in X$ to a center in $S$. Therefore, for any $i \in S$, $a^{-1}(i)$ will be all points in the cluster centered at $i$. We define
\[ \mathrm{cost}(S,a) := \sum_{i \in X} d(i, a(i)) . \]

We test this hierarchical framework for both the multiswap and supermodular local search algorithms above.

\label{submission}
